### 2016-02-03 plos_revision/coverLetter.tex

\documentclass[12pt]{letterhead_stat}
\usepackage{graphicx}
\special{papersize=8.5in,11in}
\textwidth=6.5in
\textheight=8.5in
\parindent=0.0cm
\parskip=0.4cm
\oddsidemargin=0.0in
\evensidemargin=0.0in
\topmargin=-0.594in
\begin{document}
\telephone{(510) 642-6190\\
E-mail: paciorek@stat.berkeley.edu\\
fax: (510) 642-7892}
\name{Christopher Paciorek}
\signature{\vspace{-2em}\hspace{-.3cm}\includegraphics[height=.6in,width=2in]{goodSignature.jpg}\\Christopher Paciorek\\Associate Research Statistician\\Department of Statistics\\University of California, Berkeley}
\begin{letter}{\mbox{}}
\opening{Dear Editor,}

We are responding to reviewer comments for PONE-D-15-41521.

Thank you for your consideration of our manuscript.

\closing{Sincerely,}
%\encl{names, bio}
\end{letter}
\end{document}